---
title: "Using Google Drive API with R"
output: github_document
---

## Introduction

This document aims to introduce how to use [Google Drive API](https://developers.google.com/drive/api) with R. Readers should first [create a service account](https://cloud.google.com/docs/authentication/production) and acquire the JSON key for authentication before reading this document. Below will cover how to authenticate using the service account JSON via R and demonstrate the download and upload methods. 

## Authentication

```{r}
# https://developers.google.com/drive/api/v3/reference/files/get
scopes <- c(
  "https://www.googleapis.com/auth/drive",
  "https://www.googleapis.com/auth/drive.file",
  "https://www.googleapis.com/auth/drive.readonly",
  "https://www.googleapis.com/auth/drive.metadata.readonly",
  "https://www.googleapis.com/auth/drive.appdata",
  "https://www.googleapis.com/auth/drive.metadata",
  "https://www.googleapis.com/auth/drive.photos.readonly"
)

endpoints <- httr::oauth_endpoints("google")
secrets <- jsonlite::fromJSON("./serviceAccountKey.json")

# https://www.rdocumentation.org/packages/httr/versions/1.4.2/topics/oauth_service_token
oauth_token <- httr::oauth_service_token(
  endpoint = endpoints,
  secrets = secrets,
  scope = scopes
)
```

## Download Files

Access the file without authentication. 

```{r}
fileId <- "1_zeJqQP8umrTk-evSAt3wCLxAkTKo0lC"
url <- sprintf("https://www.googleapis.com/drive/v3/files/%s?alt=media", fileId)
httr::GET(url)
```

Access the file with authentication. 

```{r}
httr::GET(url, oauth_token)
```

Saving the file. 

```{r}
httr::GET(url, oauth_token, httr::write_disk("/tmp/tmp.zip", overwrite = TRUE))
```

## Uploading Files

To upload files to Google Drive via the API, the `POST` method will be used. The request body is put in a `metadata.json` as seen below. The keys can be found from the [API document](https://developers.google.com/drive/api/v3/reference/files/create).

```{bash}
cat metadata.json
```

The script below uploads the downloaded zip file to another folder in Google Drive. 

```{r}
# https://stackoverflow.com/questions/31080363/how-to-post-multipart-related-content-with-httr-for-google-drive-api
url <- "https://www.googleapis.com/upload/drive/v3/files"

body <- list(
  metadata = httr::upload_file("./metadata.json", type = "application/json; charset=UTF-8"),
  media = httr::upload_file("/tmp/tmp.zip", type = "application/x-zip-compressed")
)

httr::POST(
  url = url,
  body = body,
  oauth_token
)
```





